### 2014-11-01 contract-language.tex
edits

An important aspect of our contract language is its ability to capture
store-level consistency guarantees, along with application-level
consistency requirements. Similar to~\cite{Burckhardt2014}, we can
rigorously define a wide variety of store semantics including those
that combine any subset of session and causality guarantees, and
multiple consistency levels.  However, for our purposes, we identify
three particular consistency levels -- eventual, causal, and strong,
commonly offered by many distributed stores with tunable consistency,
with increasing overhead in terms of latency and availability. For
each of the these three consistency levels, we capture the semantics
as a formula in our contract language, and informally describe the kind
of application-level consistency requirements that can met under the
consistency level:

@@

\item \textbf{Causal consistency}: Causally consistent operations are required
	to see a causally consistent snapshot of the object state, including the
	actions performed on the same session.  The latter requirement entails that
	if two operations $o_1$ and $o_2$ from the same session are applied to two
	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
	until the effect of $o_1$ is merged with $o_2$ in both $r_1$ and $r_2$. The
	\cf{getBalance} operation requires causal consistency, as it requires the
	operations from the same session to be visible, which cannot be guaranteed
	under just eventual consistency. We assume that causality is only tracked
	through operations on the same object. The corresponding store semantics is
	captured by the contract $\ccc$ defined below:

@@

Our goal is to map application-level consistency constraints on
operations to appropriate store-level consistency guarantees capable
of satisfying these constraints.  The ability to express both of them as
contracts in our contract language lets us compare and determine if
contract ($\cv_{op}$) of an operation ($\mathit{op}$) is weak enough to be
satisfied under a store consistency level identified by the contract
$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
relation for our contract language as following:

\begin{definition}
A contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
\begin{center}
\end{center}
\end{definition}

\vspace{-1em}
\noindent The $\Delta$ referred in the above defintion is a
conjunction of assumptions about the nature of primitive relations,
such as $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
relation $\hbZ$ is acyclic. A \emph{well-formed} axiomatic execution
($\E$) is expected to satisfy these assumptions (i.e., $\E \models
\Delta$).

If the contract ($\cv_{op}$) of an operation ($\mathit{op}$) is \emph{weaker
than} a store contract ($\cv_{st}$), then constraints expressed by the
former are implied by guarantees provided by the later. Due to the
completeness of first-order logic\cite{completeness}, this means that
any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
$\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
\cv_{op}$). Consequently, it is safe to execute operation $\mathit{op}$ under
store consistency level captured by $\cv_{st}$.

Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves
totally ordered with respect to the $\le$ relation: $\ecc \le \ccc \le
\scc$.  This concurs with the intuition that any contract satisfiable
under $\ecc$ or $\ccc$ is satisfiable under $\scc$, and any contract
that is satisfiable under $\ecc$ is satisfiable under $\ccc$. However,
we are interested in the \emph{weakest} guarantee (among $\ecc$,
$\ccc$, and $\scc$) required to satisfy the contract. We define the
corresponding consistency level as the \emph{consistency class of the
contract}. The classification scheme, presented formally in
Figure~\ref{sem:classify}, defines rules to judge consistency class of
a contact. For example, the scheme classifies \cf{getBalance} contract
($\cv_{gb}$) from Sec.~\ref{sec:motivation} as a {\sf
CausallyConsistent} contract, because the sequent $\Delta \vdash \ccc
\Rightarrow \cv_{gb}$ is valid in first-order logic (therefore,
$\cv_{gb} \le \ccc$), whereas the sequent $\Delta \vdash \ecc
\Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$).
Due to the confinement of our contract language to a decidable subset
of the logic, validity of such sequents can be decided mechanically
allowing us to automate the classification scheme in \name.

@@

In this section, we present a meta-theoretic result that certifies the
soundness of classification-based contract enforcement. To help us
state the result, we present an axiomatic model of the our system
described informally in Sec.~\ref{sec:sysmod}:

@@

\begin{theorem}[Soundness of Contract Enforcement]
\label{lem:core-preservation}
Let $\cv$ be a well-formed contract of a replicated data type operation $\mathit{op}$,
and let $\tau$ denote the consistency class of $\cv$ as determined by
the contract classification scheme. Forall well-formed execution
states $\E$, $\E'$ such that
$\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
 {\E';\langle \sigma \rangle \pll \Sigma}$
if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
\end{theorem}

The theorem states that if a data store correctly enforces $\scc$, $\ccc$,
and $\ecc$ contracts in all well-formed executions, then the same store,
extended with classification scheme, can enforce all well-formed \name
contracts. The proof of the theorem has been relegated to our tech report
and is provided as part of the supplementary
material~\cite{techrep}\footnote{The supplementary material and tech report
  also defines an operational semantics that concretizes the rewrite
  relation ($\xrightarrow{}$), along with the proof that it correctly
  enforces all \name contracts.}.

>>>>

An important aspect of our contract language is its ability to capture
store-level consistency guarantees, along with application-level
consistency requirements. Similar to~\cite{Burckhardt2014}, we can
rigorously define a wide variety of store semantics including those
that combine any subset of session and causality guarantees, and
multiple consistency levels.  However, for our purposes, we identify
three particular consistency levels -- eventual, causal, and strong,
commonly offered by many distributed stores with tunable consistency,
with increasing overhead in terms of latency and availability. For
each of the these three consistency levels, we capture the semantics
as a formula in our contract language, and informally describe the kind
of application-level consistency requirements that can be met under the
consistency level:

@@

\item \textbf{Causal consistency}: Causally consistent operations are required
	to see a causally consistent snapshot of the object state, including the
	actions performed on the same session.  The latter requirement implies that
	if two operations $o_1$ and $o_2$ from the same session are applied to two
	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
	until the effect of $o_1$ is merged with $o_2$ in both $r_1$ and $r_2$. The
	\cf{getBalance} operation requires causal consistency, as it requires the
	operations from the same session to be visible, which cannot be guaranteed
	under just eventual consistency. We assume that causality is only tracked
	through operations on the same object. The corresponding store semantics is
	captured by the contract $\ccc$ defined below:

@@

Our goal is to map application-level consistency constraints on operations
to appropriate store-level consistency guarantees capable of satisfying
these constraints.  The ability to express both these kinds of constraints
as contracts in our contract language lets us compare and determine if
contract ($\cv_{op}$) of an operation ($\mathit{op}$) is weak enough to be
satisfied under a store consistency level identified by the contract
$\cv_{st}$. Towards this end, we define a binary \emph{weaker than} relation
for our contract language as following:
\begin{definition}
A contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
\begin{center}
\end{center}
\end{definition}
\vspace{-2em}
\noindent The $\Delta$ referred in the above defintion is a
conjunction of assumptions about the nature of primitive relations,
e.g., $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
relation $\hbZ$ is acyclic. A \emph{well-formed} axiomatic execution
($\E$) is expected to satisfy these assumptions (i.e., $\E \models
\Delta$).

If the contract ($\cv_{op}$) of an operation ($\mathit{op}$) is \emph{weaker
than} a store contract ($\cv_{st}$), then constraints expressed by the
former are implied by guarantees provided by the latter.  The
completeness of first-order logic\cite{completeness} allows us to assert
that any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
$\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
\cv_{op}$). Consequently, it is safe to execute operation $\mathit{op}$ under
a store consistency level captured by $\cv_{st}$.

Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves
totally ordered with respect to the $\le$ relation: $\ecc \le \ccc \le
\scc$.  This concurs with the intuition that any contract satisfiable
under $\ecc$ or $\ccc$ is satisfiable under $\scc$, and any contract
that is satisfiable under $\ecc$ is satisfiable under $\ccc$. However,
we are interested in the \emph{weakest} guarantee (among $\ecc$,
$\ccc$, and $\scc$) required to satisfy the contract. We define the
corresponding consistency level as the \emph{consistency class of the
contract}. The classification scheme, presented formally in
Figure~\ref{sem:classify}, defines rules to judge the consistency class of
a contact. For example, the scheme classifies the \cf{getBalance} contract
($\cv_{gb}$) from Section~\ref{sec:motivation} as a {\sf
CausallyConsistent} contract, because the sequent $\Delta \vdash \ccc
\Rightarrow \cv_{gb}$ is valid in first-order logic (therefore,
$\cv_{gb} \le \ccc$), whereas the sequent $\Delta \vdash \ecc
\Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$).
Due to the confinement of our contract language to a decidable subset
of the logic, validity of such sequents can be decided mechanically
allowing us to automate the classification scheme in \name.

@@

We now present a meta-theoretic result that certifies the soundness of
classification-based contract enforcement. To help us state the result, we
define an axiomatic model of the our system described informally in
Section~\ref{sec:sysmod}:

@@

\begin{theorem}[Soundness of Contract Enforcement]
\label{lem:core-preservation}
Let $\cv$ be a well-formed contract of a replicated data type operation $\mathit{op}$,
and let $\tau$ denote the consistency class of $\cv$ as determined by
the contract classification scheme. For all well-formed execution
states $\E$, $\E'$ such that
$\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
 {\E';\langle \sigma \rangle \pll \Sigma}$,
if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
\end{theorem}

\noindent The theorem states that if a data store correctly enforces $\scc$,
$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
store, extended with the classification scheme shown in
Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
proof of the theorem is given in a tech report and is provided
as part of the supplementary material~\cite{techrep}\footnote{The
  supplementary material and tech report also defines an operational
  semantics that concretizes the rewrite relation ($\xrightarrow{}$), along
  with the proof that it correctly enforces all \name contracts.}.

### 2014-11-01 implementation.tex
edits

Performing summarization in the backing store is more complicated; the whole
process needs to be atomic from a clients perspective, but Cassandra does not
provide multi-row transactions. We have implemented a highly available
implementation of summarization that permits concurrent client operations, but
nonetheless prohibits these operations from witnessing intermediate states of
the summarization process.

>>>>

Performing summarization in the backing store is more complicated since the whole
process needs to be atomic from a client's perspective, but Cassandra does not
provide multi-row transactions. We have implemented a highly available
implementation of summarization that permits concurrent client operations, but
nonetheless prohibits these operations from witnessing intermediate states of
the summarization process.
